---
title: "5-15-19 MERS data"
author: "Megan Beaudry"
date: "May 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=F}
##this first one is a setup code chunck, ECHO = true means you will include
####all the code for the people
##ECHO is to hide or show all the code
#it will always default to waht you see at the top
##after you make a data chunck,  you need to name the data chunk with each round of R code
```




#**INTRODUCTION**
A data set on Middle East Respiratory Sydrome Corona Virus (MERS-CoV), and that has been made available by Andrw Rambaut, was utilized to practice technqiues in R studio. As such, this document contains various practce plots. 

#**MATERIALS AND METHODS**
This project used R-studio for statistical analysis. 

#STEP ONE: Load Libaries
```{r load libraries}
library(lubridate)
library(ggplot2)
library(plotly)
```


##STEP TWO: Functions
No functions were used in this analysis of the MERS-CoV data.


##STEP THREE: Data Loading
```{r load the data}
setwd('C:/Users/msb75462/Desktop/mers')
mers<-read.csv('cases.csv')
```



#STEP FOUR: Data Analysis
```{r Renaming and Formatting Table}

mers$hospitalized[890]<-c('2015-02-20')

mers<-mers[-471,]

mers$hospitalized2<-ymd(mers$hospitalized)

mers$onset2<-ymd(mers$onset)

class(mers$onset2)

day0<-min(na.omit(mers$onset2))

mers$epi.day<-as.numeric(mers$onset2-day0)
```

Endimic curves
```{r Endimic Curve by Country}
ggplot(data=mers) +
  geom_bar(mapping=aes(x=epi.day, fill=country)) +
  labs(x='Epidemic day', y='Case count', title='Global count of MERS cases by date of symptom onset',
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```
This plots shows us the Global count of MERS cases by date with a legend of countries on the side.


```{r Endimic Curve by Gender}
ggplot(data=mers) +
  geom_bar(mapping=aes(x=epi.day, fill=gender)) +
  labs(x='Epidemic day', y='Case count', title='Global count of MERS cases by date of symptom onset',
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```
This plots shows us the Global count of MERS cases by date with a legend coordiateing to the male and female genders on the side.


Univariate Plot

```{r Raw Infectious Period}
mers$infectious.period <- mers$hospitalized2-mers$onset2
```
This is the raw infectious period. 


Histogram

```{r Histogram}
class(mers$infectious.period)
mers$infectious.period <- as.numeric(mers$infectious.period, units = "days")
ggplot(data=mers) +geom_histogram(aes(x=infectious.period)) +
  labs(x='Infectious period', y='Frequency', title='Distribution of calculated MERS infectious period',
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```


Nonsocimal Infection
```{r nonsocimal infection}
mers$infectious.period2 <- ifelse(mers$infectious.period<0,0,mers$infectious.period)
ggplot(data=mers) + geom_histogram(aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency', title='Distribution of calculated MERS infectious period',
      caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```
This is an updated infectious period as this is a nonsocimal infection.

Density Plot
```{r}
ggplot(data=mers) +
  geom_density(mapping=aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency',
       title='Probability density for MERS infectious period (positive values only)', caption ="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```


       
Area Plot
```{r Area Plot}
ggplot(data=mers) +
  geom_area(stat='bin', mapping=aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency',
       title='Area plot for MERS infectious period (positive values only)', caption ="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```



Bivariate Plot

```{r Bivariate Plot}
ggplot(data=mers, mapping=aes(x=epi.day, y=infectious.period2)) +
  geom_smooth()+
  geom_point() +
  labs(x='Epidemic day', y='Infectious period',
       title='MERS infectious period (positive values only) over time', caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```



Faceting - Multipanel Plot

```{r Faceting}
ggplot(data=mers, mapping=aes(x=epi.day, y=infectious.period2)) +
  geom_point(mapping = aes(color=country)) +
  facet_wrap(~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x='Epidemic day', y='Infectious period',
  title='MERS infectious period (positive values only) over time', caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

Seperate Plots for Male and Female 

```{r facet male and female}
ggplot(data=subset(mers, gender %in% c('M', 'F') &
                     country %in% c('KSA', 'Oman', 'Iran', 'Jordan', 'Qatar', 'South Korea','UAE')),
       mapping=aes(x=epi.day, y=infectious.period2)) +
  geom_point(mapping = aes(color=country)) +
  facet_grid(gender ~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x='Epidemic day', y='Infectious period',
       title='MERS infectious period by gender and country',
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

Interactive Plot
```{r PLotly}
epi.curve <- ggplot(data=mers) +
geom_bar(mapping=aes(x=epi.day)) +
labs(x='Epidemic day', y='Case count', title='Global count of MERS cases by date of symptom onset',
caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
ggplotly(epi.curve)
```


#**Results**
This study reflected.........

#**Conclusion**

